Synthesizing Verified Components for Cyber Assured Systems Engineering \thanks{DISTRIBUTION STATEMENT A. Approved for public release.}

### 2022-03-27 paper/testing.tex
Light proofreading

Before discussing the final steps in synthesis, we take a moment to discuss how code-contracts, and any AGREE specification in general, can be traditionally tested to build assurance.
Writing and proving properties about AGREE specifications is hard not just because it is a formal process but  because it reasons about constraints that define the entire input space for a component, and the entire output space, at once.
The case studies showed that engineers, and formal method practitioners, make mistakes writing these specifications especially if the computation or temporal reasoning is complex.

@@

AGREE already detects, and reports, contradicting and inconsistent contracts, but it does not check for vacuity and under-specification.
The tempting approach to check these issues is to use the synthesize path and then apply traditional testing.
Such an approach adds extra steps that are not needed as it is possible to bridge the formal model in AGREE with its proof system to traditional reasoning with test.

Unit tests are created in AGREE by writing a system specification that assumes the test input and then guarantees the test output.
The system is then implemented with the component under test and its specification.
The AGREE analysis then proves if the component guarantees, given the test inputs, are strong enough to prove the guaranteed test outputs.
Here the system assumptions strengthen the component assumptions to a single input, and the system guarantees weaken, or leave unchanged, the component output for that test input, depending on the goal of the test.
AGREE proves the component implements the test, and the Liskov substitution principle holds.

@@

\figref{fig:test} is a unit test for the high-assurance monitor in \figref{fig:hardened}.
The test checks if the monitor output coincides with the response that comes one step after the request.
It also checks that the alarm does not fire.
\lineref{line:tst-index} creates an index to use to create the test input.
\lineref{line:tst-assume-1} defines when the response appears relative to the index.
\lineref{line:tst-assume-2} defines the request behavior.
\linesref{line:tst-guarantee-1}{line:tst-guarantee-2} define the expected test output.
The test additionally proves that nothing happens after the initial test input and output (e.g., no more input is provided and no more output is generated).
AGREE proves the test correct.
\linesref{line:tst-imp-start}{line:tst-imp-end} implement the system using the component (\lineref{line:tst-imp-comp}).
\figref{fig:test-output} is the AGREE output for the test is showing that it passes.

>>>>

Before discussing the final steps in synthesis, we take a moment to
discuss how code contracts, and any AGREE specification in general, can be traditionally tested to build assurance.
Writing and proving properties about AGREE specifications is hard not just because it is a formal process but  because it reasons about constraints that define the entire input space for a component, and the entire output space, at once.
The case studies showed that engineers, and formal method practitioners, make mistakes writing these specifications especially if the computation or temporal reasoning is complex.

@@

AGREE already detects, and reports, contradicting and inconsistent contracts, but it does not check for vacuity and under-specification.
The tempting approach to check these issues is to use the synthesis path and then apply traditional testing.
Such an approach adds extra steps that are not needed, as it is
possible to bridge the formal model in AGREE, with its proof system, to
traditional validation using tests.

Unit tests are created in AGREE by writing a system specification that assumes the test input and then guarantees the test output.
The system is then implemented with the component under test and its specification.
The AGREE analysis then proves if the component guarantees, given the test inputs, are strong enough to prove the guaranteed test outputs.
Here the system assumptions strengthen the component assumptions to a single input, and the system guarantees weaken, or leave unchanged, the component output for that test input, depending on the goal of the test.
AGREE proves that the component implements the test, and the Liskov substitution principle holds.

@@

\figref{fig:test} is a unit test for the high-assurance monitor in \figref{fig:hardened}.
The test checks if the monitor output coincides with the response that comes one step after the request.
It also checks that the alarm does not fire.
\lineref{line:tst-index} creates an index to use to create the test input.
\lineref{line:tst-assume-1} defines when the response appears relative to the index.
\lineref{line:tst-assume-2} defines the request behavior.
\linesref{line:tst-guarantee-1}{line:tst-guarantee-2} define the expected test output.
The test additionally proves that nothing happens after the initial test input and output (e.g., no more input is provided and no more output is generated).
AGREE proves the test correct.
\linesref{line:tst-imp-start}{line:tst-imp-end} implement the system using the component (\lineref{line:tst-imp-comp}).
\figref{fig:test-output} is the AGREE output for the test, showing that it passes.
